HyperLogLog: Exponentially Bad in Adversarial Settings

### 2020-10-13 docs/HLL_exp_bad/main.tex
more tweakery

Finally, we conclude our analysis by reiterating the key point: we have shown that no efficient S4 adversary ${\cal A}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage. This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$. It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries. For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.

@@

This construction would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations. The exact slowdown relative to, say, Murmur3, depends on platform specifics, but can be expected to be in the range 2-fold to 5-fold.

@@

@@

It is an important open problem to close the performance gap that exists between fast but non-cryptographic hashing used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security in the face of adversarial inputs.

We also note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL filters. The reason is that one can no longer emulate running a single filter by taking the maximum over all buckets with each index $j$ in the different filters, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.

@@

We have given a formal analysis of using a secretly-keyed PRF as a countermeasure to our attacks. The reliance on secret keys means sacrificing the mergeability property of HLL sketches in distributed environments (where the keys in the different HLLs are independently generated). It is an interesting cryptographic challenge to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability. This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.

>>>>

Finally, we conclude our analysis by reiterating the key point: we have shown that no efficient S4 adversary ${\cal A}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage. This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$. It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries. For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL. The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.

@@

@@

These constructions would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations. It would be a simple exercise to add the above PRF proposals to an existing HLL implementation and benchmark them against, say, Murmur3.  The exact slowdown depends on platform specifics, but based on microbenchmarks can be expected to be in the range 2-fold to 5-fold compared to Murmur3.

@@

We note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL filters. The reason is that one can no longer emulate running a single filter by taking the maximum over all buckets with each index $j$ in the different filters, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.

@@

We have given a formal analysis of using a secretly-keyed PRF as a countermeasure to our attacks. The reliance on secret keys means sacrificing the mergeability property of HLL sketches in distributed environments (where the keys in the different HLLs are independently generated). It is an interesting cryptographic challenge to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability. This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security. It is also an important open problem to close the performance gap that exists between fast but non-cryptographic hashing used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security in the face of adversarial inputs.
